---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# moodef

<!-- badges: start -->
<!-- badges: end -->

The goal of the `moodef` package is to support the definition of [*Moodle*](https://moodle.org/) elements taking advantage of the power that R offers. In particular, in this first version, it allows the definition of questions to be included in the question bank to define quizzes.

To define the questions for the quizzes we can use the component for this purpose that includes *Moodle*, based on entering data through screens. It allows the import and export of questions in various formats, including xml.

Complementary, using the `moodef` package, we can define questionnaires from R. We have generalized 9 types of questions and simplified their definition. So, we define a question by calling a function or simply including a row in a data frame or a csv file. The result is an xml file that we import into *Moodle*. If necessary, some parameter not considered in the generalization can be defined or adjusted there.

Although not shown in the next example, in each question we can include an image that is embedded in xml. We can set up the size of the images so that they are homogeneous when displayed in quizzes.

In addition to facilitating the definition of questions manually, the infrastructure offered by this package can be used to generate questions automatically or semi-automatically from R.

## Installation

You can install the development version of `moodef` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("josesamos/moodef")
```

## Example

This is a basic example which shows the definition of a question using the function:

```{r example}
library(moodef)

qc <- question_category(category = 'Initial test',
                        copyright = 'Copyright © 2024 Universidad de Granada',
                        license = 'License Creative Commons Attribution-ShareAlike 4.0') |>
  define_question(
    question = 'What are the basic arithmetic operations?',
    answer = 'Addition, subtraction, multiplication and division.',
    a_1 = 'Addition and subtraction.',
    a_2 = 'Addition, subtraction, multiplication, division and square root.'
  )

file <- tempfile(fileext = '.xml')
qc <- qc |>
  generate_xml_file(file)
```

First, we create an object using the `question_category` function and configure general aspects of the definition in it. Next, we define the questions, as many as we need, using the `define_question` function (the type of the questions is deduced from the definition). Finally, we generate the questions in xml format, in the form of a string or file. We show the result below.

````{verbatim, lang = "xml"}
<?xml version="1.0" encoding="UTF-8"?>
<quiz>
  <question type="category">
    <category> <text>$course$/top/Initial test</text> </category>
    <info format="html"> <text></text> </info>
    <idnumber></idnumber>
  </question>
  <question type="multichoice">
<name> <text>q_001_multichoice_what_are_the_basic_arithmetic_operations</text> </name>
<questiontext format="html">
  <text><![CDATA[
     <!-- Copyright © 2024 Universidad de Granada -->
     <!-- License Creative Commons Attribution-ShareAlike 4.0 -->
     <p>What are the basic arithmetic operations?</p>]]></text>
     
</questiontext>
<generalfeedback format="html"> <text></text> </generalfeedback>
<defaultgrade>1.0000000</defaultgrade>
<penalty>0.5</penalty>
<hidden>0</hidden>
<idnumber></idnumber>
<single>true</single>
<shuffleanswers>true</shuffleanswers>
<answernumbering>abc</answernumbering>
<showstandardinstruction>0</showstandardinstruction>
<correctfeedback format="moodle_auto_format"> <text>Correct.</text> </correctfeedback>
<partiallycorrectfeedback format="moodle_auto_format"> <text></text> </partiallycorrectfeedback>
<incorrectfeedback format="moodle_auto_format"> <text>Incorrect.</text> </incorrectfeedback>
<answer fraction="100" format="html">
   <text>Addition, subtraction, multiplication and division.</text>
   <feedback format="html"> <text>Correct.</text> </feedback>
</answer>
<answer fraction="-50.000000000000000" format="html">
   <text>Addition and subtraction.</text>
   <feedback format="html"> <text>Incorrect.</text> </feedback>
</answer>
<answer fraction="-50.000000000000000" format="html">
   <text>Addition, subtraction, multiplication, division and square root.</text>
   <feedback format="html"> <text>Incorrect.</text> </feedback>
</answer>
</question>
</quiz>
````

We can call the `define_question` function for each question to be defined, with the same parameters. Alternatively, we can create a data frame or a csv file (with the functions available in the package) and include in them a row for each question, a column for each parameter. Below is the content of a csv file.

````{verbatim, lang = "csv"}
"type","question","image","image_alt","answer",,,
,"What are the basic arithmetic operations?",,,"Addition, subtraction, multiplication and division.","Addition and subtraction.","Addition, subtraction, multiplication, division and square root.",
,"Match each operation with its symbol.",,,"Addition<|>+","Subtraction<|>-","Multiplication<|>*",
,"The square root is a basic arithmetic operation.",,,"False",,,
,"What basic operation does it have as a ""+"" symbol?",,,"Addition",,,
,"The symbol for addition is [[1]], the symbol for subtraction is [[2]].",,,"+","-",,
"x","The symbol for addition is [[1]], the symbol for subtraction is [[2]].",,,"+","-",,
"h","Sort the result from smallest to largest.",,,"6/2","6-2","6+2","6*2"
"x","Sort the result from smallest to largest.",,,"6/2","6-2","6+2","6*2"
,"What is the result of SQRT(4)?",,,"2","-2",,
,"What is the result of 4/3?",,,"1.33<|>0.03",,,
,"Describe the addition operation.",,,,,,
````

The generation of the questions would be similar, as shown below.

```{r csv}
file <- system.file("extdata", "questions.csv", package = "moodef")
qc <- question_category(category = 'Initial test',
                        copyright = 'Copyright © 2024 Universidad de Granada',
                        license = 'License Creative Commons Attribution-ShareAlike 4.0') |>
  define_questions_from_csv(file = file)

file <- tempfile(fileext = '.xml')
qc <- qc |>
  generate_xml_file(file)
```

We do not show the content of the xml file obtained because in that example there are several questions defined.
